---
title: "Final Project"
author: "Marco Sciorilli"
date: "17/12/2020"
output: html_document
    
institute: University of Trieste
subtitle: Statical Methods for Data Science
fontsize: 10pt
---


Perform some explanatory analysis for your data, especially by use of graphical tools.
```{r setup, include=FALSE}
library(DAAG)
library(MASS)
data<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-regioni/dpc-covid19-ita-regioni.csv")
```

```{r setup, include=FALSE}
dataupdated<-subset(data, data >= "2020-09-01T17:00:00")
dataupdatedLombardia<-subset(dataupdated, codice_regione == "3")
dataupdatedLombardiacleaned <-subset(dataupdatedLombardia, select = -c(stato,codice_regione,lat,long,denominazione_regione,note,note_test,note_casi,codice_nuts_1,codice_nuts_2,data) )
dataupdatedLombardiacleaned$giorno<- seq.int(nrow(dataupdatedLombardiacleaned))
dataupdatedLombardiacleaned$giorno_settimana<-seq.int(nrow(dataupdatedLombardiacleaned))
dataupdatedLombardiacleaned$picco<-seq.int(nrow(dataupdatedLombardiacleaned))
for (i in 7:length(dataupdatedLombardia[,1])) {
  if(i%%7==0){
dataupdatedLombardiacleaned$giorno_settimana[i]<-"martedi"
  }
  if(i%%7==1){
dataupdatedLombardiacleaned$giorno_settimana[i]<-"mercoledi"  }
  if(i%%7==2){
dataupdatedLombardiacleaned$giorno_settimana[i]<-"giovedi"  }
  if(i%%7==3){
dataupdatedLombardiacleaned$giorno_settimana[i]<-"venerdi"  }
  if(i%%7==4){
dataupdatedLombardiacleaned$giorno_settimana[i]<-"sabato" }
  if(i%%7==5){
dataupdatedLombardiacleaned$giorno_settimana[i]<- "domenica"}
  if(i%%7==6){
dataupdatedLombardiacleaned$giorno_settimana[i]<-"lunedi"  }

}
dataupdatedLombardiacleaned$giorno_settimana[1]<-"martedi"
dataupdatedLombardiacleaned$giorno_settimana[2]<-"mercoledi"
dataupdatedLombardiacleaned$giorno_settimana[3]<-"giovedi"
dataupdatedLombardiacleaned$giorno_settimana[4]<-"venerdi"
dataupdatedLombardiacleaned$giorno_settimana[5]<-"sabato"
dataupdatedLombardiacleaned$giorno_settimana[6]<- "domenica"
dataupdatedLombardiacleaned$giorno_settimana[7]<-"lunedi"
for (i in 1:35){      #83 punto più credibile
dataupdatedLombardiacleaned$picco[i]<-"pre-picco"
}
for (i in 36:83){ #83 punto più credibile
dataupdatedLombardiacleaned$picco[i]<-"picco"
}
for (i in 84:length(dataupdatedLombardia[,1])) {
dataupdatedLombardiacleaned$picco[i]<-"post-picco"
}
dulc<-dataupdatedLombardiacleaned
```

```{r}
library("ggpubr")
library(ggplot2)
for (i in 1:20) {  
      file_name = paste("untouched", i, ".png", sep="")
  png(file=file_name,
width=6, height=4, units="in", res=300)
#print(cor.test(dulc[,i],dulc[,7], method=c("spearman")))
print(ggplot(dulc,  aes(dulc[,i],dulc[,5])) + 
  geom_point(aes()) +
  theme_minimal()+
  theme(legend.position = "top")+
  xlab(colnames(dulc[i]))+
  ylab(colnames(dulc[5])))
dev.off()
}

```


```{r}
dataupdatedLombardiacleaned <-subset(dataupdatedLombardiacleaned, select = -c(tamponi_test_antigenico_rapido,tamponi_test_molecolare,totale_positivi_test_antigenico_rapido,totale_positivi_test_molecolare, casi_da_screening, casi_da_sospetto_diagnostico, ingressi_terapia_intensiva) )
dulc<-dataupdatedLombardiacleaned
dulcnonlog<-dataupdatedLombardiacleaned

library(PerformanceAnalytics)
file_name = paste("cross", 1, ".png", sep="")
  png(file=file_name,
width=6, height=4, units="in", res=300)
chart.Correlation(dulc[,1:5],main="Esplorazione nuovi casi")
dev.off()
file_name = paste("cross", 2, ".png", sep="")
  png(file=file_name,
width=6, height=4, units="in", res=300)
chart.Correlation(dulc[,6:7],main="Esplorazione nuovi casi")
dev.off()
file_name = paste("cross", 3, ".png", sep="")
  png(file=file_name,
width=6, height=4, units="in", res=300)
chart.Correlation(dulc[,8:9],main="Esplorazione nuovi casi")
dev.off()
file_name = paste("cross", 4, ".png", sep="")
  png(file=file_name,
width=6, height=4, units="in", res=300)
chart.Correlation(dulc[,10:13],main="Esplorazione nuovi casi")
dev.off()
```



```{r}
dulc<- log(dulc[1:12])
dulc$giorno<-log( seq.int(nrow(dulc)))
dulc$giorno_settimana<-seq.int(nrow(dulc))
dulc$picco<-seq.int(nrow(dulc))
for (i in 7:length(dataupdatedLombardia[,1])) {
  if(i%%7==0){
dulc$giorno_settimana[i]<-"martedi"
  }
  if(i%%7==1){
dulc$giorno_settimana[i]<-"mercoledi"  }
  if(i%%7==2){
dulc$giorno_settimana[i]<-"giovedi"  }
  if(i%%7==3){
dulc$giorno_settimana[i]<-"venerdi"  }
  if(i%%7==4){
dulc$giorno_settimana[i]<-"sabato" }
  if(i%%7==5){
dulc$giorno_settimana[i]<- "domenica"}
  if(i%%7==6){
dulc$giorno_settimana[i]<-"lunedi"  }

}
dulc$giorno_settimana[1]<-"martedi"
dulc$giorno_settimana[2]<-"mercoledi"
dulc$giorno_settimana[3]<-"giovedi"
dulc$giorno_settimana[4]<-"venerdi"
dulc$giorno_settimana[5]<-"sabato"
dulc$giorno_settimana[6]<- "domenica"
dulc$giorno_settimana[7]<-"lunedi"
for (i in 1:35){      #83 punto più credibile
dulc$picco[i]<-"pre-picco"
}
for (i in 36:83){      #83 punto più credibile
dulc$picco[i]<-"picco"
}
for (i in 84:length(dataupdatedLombardia[,1])) {
dulc$picco[i]<-"post-picco"
}
dulc$colore<-seq.int(nrow(dulc))

for (i in 1:65){      #83 punto più credibile
dulc$colore[i]<-"bianco"
}
for (i in 66:88) {
dulc$colore[i]<-"rosso"
}
for (i in 89:103) {
dulc$colore[i]<-"arancione"
}
for (i in 104:114) {
dulc$colore[i]<-"giallo"
}
for (i in 115:118) {
dulc$colore[i]<-"rosso"
}
for (i in 119:121) {
dulc$colore[i]<-"arancione"
}
for (i in 122:125) {
dulc$colore[i]<-"rosso"
}
for (i in 126:126) {
dulc$colore[i]<-"rosso"
}
for (i in 127:128) {
dulc$colore[i]<-"rosso"
}
for (i in 129:130) {
dulc$colore[i]<-"giallo"
}
for (i in 131:137) {
dulc$colore[i]<-"arancione"
}
for (i in 138:145) {
dulc$colore[i]<-"rosso"
}

for (i in 146:153) {
dulc$colore[i]<-"arancione"
}

for (i in 144:length(dataupdatedLombardia[,1])) {
dulc$colore[i]<-"giallo"
}
for (i in 1:13) {
      file_name = paste("touched", i, ".png", sep="")
  png(file=file_name,
width=6, height=4, units="in", res=300)
#print(cor.test(dulc[,i],dulc[,7], method=c("spearman")))
print(ggplot(dulc,  aes(dulc[,i],dulc[,5])) + 
  geom_point(aes(color=as.factor(dulc[,15]))) +
  theme_minimal()+
  theme(legend.position = "top")+
  xlab(colnames(dulc[i]))+
  ylab(colnames(dulc[5])))
dev.off()
}


library(PerformanceAnalytics)
file_name = paste("crossnew", 1, ".png", sep="")
  png(file=file_name,
width=6, height=4, units="in", res=300)
chart.Correlation(dulc[,1:5],main="Esplorazione nuovi casi")
dev.off()
file_name = paste("crossnew", 2, ".png", sep="")
  png(file=file_name,
width=6, height=4, units="in", res=300)
chart.Correlation(dulc[,6:7],main="Esplorazione nuovi casi")
dev.off()
file_name = paste("crossnew", 3, ".png", sep="")
  png(file=file_name,
width=6, height=4, units="in", res=300)
chart.Correlation(dulc[,8:9],main="Esplorazione nuovi casi")
dev.off()
file_name = paste("crossnew", 4, ".png", sep="")
  png(file=file_name,
width=6, height=4, units="in", res=300)
chart.Correlation(dulc[,10:13],main="Esplorazione nuovi casi")
dev.off()
```
Da 1 settembre a 5 novembre zona bianca

Da 6 novembre a 28 novembre zona rossa 

da 29 novembre a 12 dicembre zona arancione
da 13 a 23 zona gialla
da 24 a 27 rossa 
28-29-30 arancione
31 - 3 rossa
4 arancione
5-6 rossa

7-8 gialla
9-15 arancione
16-23 zona rossa
24-1 febbraio arancione






totale_ospedalizzati
totale_positivi
variazione_totale_positivi

sono ad alto rischio di ridondanza





Describe the quality of the data and discuss whether and how a plausible statistical model could be posed.

Build a model for your response variable Y. To this aim you can adopt any among the regression techniques covered during the course. Comment the estimates from the best model obtained.

imposto un modello con tutte le variabili con andamento simile, controllo la multicollinearità

Seleziono solo la variabile con p-value minore
```{r}
dulc.lm1 <- lm(totale_positivi ~  terapia_intensiva+ricoverati_con_sintomi+totale_ospedalizzati+isolamento_domiciliare, data = dulc)
summary(dulc.lm1)
(vif(dulc.lm1))
library(stargazer)
stargazer((vif(dulc.lm1)), type = "latex", out = "vif.tex")
library(xtable)
xtable(anova(dulc.lm1))
anova(dulc.lm1)
```

```{r}
dulc.lm1.1 <- lm(totale_positivi ~  terapia_intensiva, data = dulc)
dulc.lm1.2 <- lm(totale_positivi ~  (terapia_intensiva)* picco, data = dulc)
dulc.lm1.3 <- lm((totale_positivi) ~ I((terapia_intensiva)^2)+I((terapia_intensiva)), data = dulc)
dulc.lm1.4 <- lm((totale_positivi) ~ I((terapia_intensiva)^2)*picco+I((terapia_intensiva))*picco, data = dulc)
AIC <- rbind(extractAIC(dulc.lm1.1)[2],
extractAIC(dulc.lm1.2)[2],
extractAIC(dulc.lm1.3)[2],
extractAIC(dulc.lm1.4)[2]
)
#BIC
BIC <- rbind(extractAIC(dulc.lm1.1, k = log(n))[2], 
             extractAIC(dulc.lm1.2, k = log(n))[2], 
             extractAIC(dulc.lm1.3, k = log(n))[2],
            extractAIC(dulc.lm1.4, k = log(n))[2]
             )
model <- c("dulc.lm1.1", "dulc.lm1.2", "dulc.lm1.3","dulc.lm1.4")
xtable(cbind.data.frame(model,AIC,BIC))

```

```{r}
plot( dulc[,2],dulc[,5])
#curve(predict(dulc.lm, data.frame(ricoverati_con_sintomi=x,picco=x)) ,add=TRUE,col="red")
lines(dulc[,2] , dulc.lm1.4$fitted.values, lty =1 , col = "red")
```



















```{r}
dulc.lm2 <- lm(totale_positivi ~  nuovi_positivi+ variazione_totale_positivi, data = dulc)
summary(dulc.lm2)
vif(dulc.lm2)
library(stargazer)
stargazer((vif(dulc.lm2)), type = "latex", out = "vif2.tex")
library(xtable)
xtable(anova(dulc.lm2))
anova(dulc.lm2)
dulc.lm2 <- lm(totale_positivi ~  nuovi_positivi*picco, data = dulc)
```


```{r}
plot(dulc[,7],dulc[,5])
curve((as.numeric(dulc.lm2$coefficients[1])+as.numeric(dulc.lm2$coefficients[2])*x), from=0, to=15000, add=TRUE,col="red") 
curve(as.numeric(dulc.lm2$coefficients[1])+as.numeric(dulc.lm2$coefficients[3]) + as.numeric(dulc.lm2$coefficients[5])*x+as.numeric(dulc.lm2$coefficients[2])*x, from=0, to=15000, add=TRUE,col="red") 
curve((as.numeric(dulc.lm2$coefficients[1])+as.numeric(dulc.lm2$coefficients[4])+as.numeric(dulc.lm2$coefficients[6])*x+as.numeric(dulc.lm2$coefficients[2])*x), from=0, to=15000, add=TRUE,col="red") 
#lines(dulc[,7] , dulc.lm2$fitted.values, lty =1 , col = "red")
```

```{r}
dulc.lm2.1 <- lm(totale_positivi ~  nuovi_positivi, data = dulc)
dulc.lm2.2 <- lm(totale_positivi ~  nuovi_positivi* picco, data = dulc)
dulc.lm2.3 <- lm((totale_positivi) ~ I((nuovi_positivi)^2)+I((nuovi_positivi)), data = dulc)
dulc.lm2.4 <- lm((totale_positivi) ~ I((nuovi_positivi)^2)*picco+I((nuovi_positivi))*picco, data = dulc)
AIC <- rbind(extractAIC(dulc.lm2.1)[2],
extractAIC(dulc.lm2.2)[2],
extractAIC(dulc.lm2.3)[2],
extractAIC(dulc.lm2.4)[2]
)
#BIC
BIC <- rbind(extractAIC(dulc.lm2.1, k = log(n))[2], 
             extractAIC(dulc.lm2.2, k = log(n))[2], 
             extractAIC(dulc.lm2.3, k = log(n))[2],
            extractAIC(dulc.lm2.4, k = log(n))[2]
             )
model <- c("dulc.lm2.1", "dulc.lm2.2", "dulc.lm2.3","dulc.lm2.4")
xtable(cbind.data.frame(model,AIC,BIC))

```

```{r}
dulc.lm3 <- lm(totale_positivi ~  dimessi_guariti+ deceduti, data = dulc)
summary(dulc.lm3)
vif(dulc.lm3)
library(stargazer)
stargazer((vif(dulc.lm3)), type = "latex", out = "vif2.tex")
library(xtable)
xtable(anova(dulc.lm3))
anova(dulc.lm3)
```

```{r}
dulc.lm3.1 <- lm(totale_positivi ~  dimessi_guariti, data = dulc)
dulc.lm3.2 <- lm(totale_positivi ~  dimessi_guariti* picco, data = dulc)
dulc.lm3.3 <- lm((totale_positivi) ~ I((dimessi_guariti)^2)+I((dimessi_guariti)), data = dulc)
dulc.lm3.4 <- lm((totale_positivi) ~ I((dimessi_guariti)^2)*picco+I((dimessi_guariti))*picco, data = dulc)
AIC <- rbind(extractAIC(dulc.lm3.1)[2],
extractAIC(dulc.lm3.2)[2],
extractAIC(dulc.lm3.3)[2],
extractAIC(dulc.lm3.4)[2]
)
#BIC
BIC <- rbind(extractAIC(dulc.lm3.1, k = log(n))[2], 
             extractAIC(dulc.lm3.2, k = log(n))[2], 
             extractAIC(dulc.lm3.3, k = log(n))[2],
            extractAIC(dulc.lm3.4, k = log(n))[2]
             )
model <- c("dulc.lm3.1", "dulc.lm3.2", "dulc.lm3.3","dulc.lm3.4")
xtable(cbind.data.frame(model,AIC,BIC))

```


```{r}
plot((dulc[,8]),dulc[,5])
#curve((as.numeric(dulc.lm3$coefficients[1])+as.numeric(dulc.lm3$coefficients[2])*(x)), from=2, to=3, add=TRUE,col="red") 
#curve((10+20*log(x-11)), from=0, to=15, add=TRUE,col="red") 
#curve(as.numeric(dulc.lm3$coefficients[1])+as.numeric(dulc.lm3$coefficients[3]) + as.numeric(dulc.lm3$coefficients[5])*x+as.numeric(dulc.lm3$coefficients[2])*x, from=0, to=15000, add=TRUE,col="red") 
#curve((as.numeric(dulc.lm3$coefficients[1])+as.numeric(dulc.lm3$coefficients[4])+as.numeric(dulc.lm3$coefficients[6])*x+as.numeric(dulc.lm3$coefficients[2])*x), from=0, to=0.89, add=TRUE,col="red") 
lines((dulc[,8]) , dulc.lm3.4$fitted.values, lty =1 , col = "red")
```
```{r}
dulc.lm6 <- lm(totale_positivi ~  totale_casi+tamponi+casi_testati+giorno, data = dulc)
summary(dulc.lm6)
vif(dulc.lm6)
library(stargazer)
stargazer((vif(dulc.lm6)), type = "latex", out = "vif3.tex")
library(xtable)
xtable(anova(dulc.lm6))
anova(dulc.lm6)

```
```{r}
dulc.lm4.1 <- lm(totale_positivi ~  totale_casi, data = dulc)
dulc.lm4.2 <- lm(totale_positivi ~  totale_casi* picco, data = dulc)
dulc.lm4.3 <- lm((totale_positivi) ~ I((totale_casi)^2)+I((totale_casi)), data = dulc)
dulc.lm4.4 <- lm((totale_positivi) ~ I((totale_casi)^2)*picco+I((totale_casi))*picco, data = dulc)
AIC <- rbind(extractAIC(dulc.lm4.1)[2],
extractAIC(dulc.lm4.2)[2],
extractAIC(dulc.lm4.3)[2],
extractAIC(dulc.lm4.4)[2]
)
#BIC
BIC <- rbind(extractAIC(dulc.lm4.1, k = log(n))[2], 
             extractAIC(dulc.lm4.2, k = log(n))[2], 
             extractAIC(dulc.lm4.3, k = log(n))[2],
            extractAIC(dulc.lm4.4, k = log(n))[2]
             )
model <- c("dulc.lm4.1", "dulc.lm4.2", "dulc.lm4.3","dulc.lm4.4")
xtable(cbind.data.frame(model,AIC,BIC))

```

scelgo totale casi
```{r}
plot( dulc[,10],dulc[,5])
lines(dulc[,10] , dulc.lm4$fitted.values, lty =1 , col = "red")
```



```{r}
dulc.lm5 <- lm(totale_positivi ~ I((terapia_intensiva)^2)*picco+I((terapia_intensiva))*picco+I(totale_casi^2)*picco+I(totale_casi)*picco+I(dimessi_guariti^2)*picco+I(dimessi_guariti)*picco+I((nuovi_positivi)^2)*picco+I((nuovi_positivi))*picco, data = dulc)
summary(dulc.lm5)
plot(dulc.lm5)
log(vif(dulc.lm5))
```


```{r}
plot( dulc[,13],dulc[,5])
lines(dulc[,13] , dulc.lm5$fitted.values, lty =1 , col = "red")
```


```{r}
(anova(dulc.lm5))
AIC <- rbind(extractAIC(dulc.lm5)[2]
)
#BIC
BIC <- rbind(extractAIC(dulc.lm5, k = log(n))[2]
             )
model <- c("dulc.lm5")
cbind.data.frame(model,AIC,BIC)
plot(dulc.lm5,col=as.factor(dulc[,15]))
```

```{r}
dulc.lm7 <- glm(totale_positivi ~ I((terapia_intensiva)^2)*picco+I((terapia_intensiva))*picco+I(totale_casi^2)*picco+I(totale_casi)*picco+I(dimessi_guariti^2)*picco+I(dimessi_guariti)*picco+I((nuovi_positivi)^2)*picco+I((nuovi_positivi))*picco, data =dulc[1:(length(dataupdatedLombardia[,1])-15),],family=poisson(link=log))
summary(dulc.lm7)
plot(dulc.lm7,col=as.factor(dulc[,15]))
```



```{r}
dulc.lm8 <- glm(totale_positivi ~ I((terapia_intensiva)^2)*picco+I((terapia_intensiva))*picco+I(totale_casi^2)*picco+I(totale_casi)*picco+I(dimessi_guariti^2)*picco+I(dimessi_guariti)*picco+I((nuovi_positivi)^2)*picco+I((nuovi_positivi))*picco, data = dulc, family=Gamma(link=log))
summary(dulc.lm8)
for (i in 1:4){
      file_name = paste("residual", i, ".png", sep="")
  png(file=file_name,
width=6, height=4, units="in", res=300)
plot(dulc.lm8,which=i,col=as.factor(dulc[,15]))
dev.off()
}
```
```{r}
dulc.lm7 <- glm((totale_positivi) ~ log(terapia_intensiva)+I(totale_casi^2)*picco+I(totale_casi)*picco+I(dimessi_guariti^2)*picco+I(dimessi_guariti)*picco+nuovi_positivi*colore, data =dulc[1:(length(dataupdatedLombardia[,1])-15),],family=poisson(link=log))
summary(dulc.lm7)
plot(dulc.lm7,col=as.factor(dulc[,15]))

```



```{r}
library(lasso2)
dulc.lm8 <- l1ce(totale_positivi ~ log(isolamento_domiciliare)+I(totale_casi^2)*picco+I(totale_casi)*picco+I(dimessi_guariti^2)*picco+I(dimessi_guariti)*picco+terapia_intensiva* picco, data = dulc)
summary(dulc.lm8)$coefficients
plot(dulc.lm8$fitted,dulc.lm8$residuals,col=as.factor(dulc[,15]))
library("car")
qqPlot(dulc.lm8$residuals,lwd=1)
```

```{r}
plot( dulc[,13],dulc[,5])
lines(dulc[,13] , dulc.lm8$fitted.values, lty =1 , col = "red")
```
```{r}
library(mgcv)
dulc.lm9 <- gam(exp(totale_positivi) ~ s(log(isolamento_domiciliare))+s(poly(totale_casi, degree = 2, raw = TRUE)*picco)+s(poly(dimessi_guariti, degree = 2, raw = TRUE)*picco)+ s(terapia_intensiva* picco), data = dulc, family=quasipoisson(link=log))
summary(dulc.lm7)
plot(dulc.lm9,col=as.factor(dulc[,15]))
```

```{r}
plot( dulc[,13],dulc[,5])
lines(dulc[,13] , dulc.lm9$fitted.values, lty =1 , col = "red")
```


```{r}
coverage_log <- exp(predict(dulc.lm7, interval ="confidence"))
exp(predict(dulc.lm7, interval ="confidence"))[1:14,]
```


```{r}
library(randomForest)
dulcrf<-subset(dulc, select = -c(variazione_totale_positivi) )
dulc.lm10 <- randomForest(totale_positivi ~ ., data=dulcrf[1:(length(dataupdatedLombardia[,1])-15),], importance=TRUE) 
print(dulc.lm10)
importance(dulc.lm10)
random_forest<-predict(dulc.lm10, dulcrf[,])

random_forest
```
```{r}

is.dom.gam<-randomForest(nuovi_positivi ~ ., data=dulcrf[1:(length(dataupdatedLombardia[,1])-15),], importance=TRUE)



tot.cas.gam<-randomForest(totale_casi ~ ., data=dulcrf[1:(length(dataupdatedLombardia[,1])-15),], importance=TRUE)

dim.guar.gam<-randomForest(dimessi_guariti ~ ., data=dulcrf[1:(length(dataupdatedLombardia[,1])-15),], importance=TRUE)


ter.int.gam<-randomForest(terapia_intensiva ~ ., data=dulcrf[1:(length(dataupdatedLombardia[,1])-15),], importance=TRUE)


```



```{R}
dulc.lm8 <- glm(totale_positivi ~ I((terapia_intensiva)^2)*picco+I((terapia_intensiva))*picco+I(totale_casi^2)*picco+I(totale_casi)*picco+I(dimessi_guariti^2)*picco+I(dimessi_guariti)*picco+I((nuovi_positivi)^2)*picco+I((nuovi_positivi))*picco, data = dulc[1:(length(dataupdatedLombardia[,1])-15),], family=Gamma(link=log))
predictions<-predict.glm(dulc.lm8, data.frame(terapia_intensiva=predict(ter.int.gam, dulc[(length(dataupdatedLombardia[,1])-15):length(dataupdatedLombardia[,1]),]),dimessi_guariti=predict(dim.guar.gam, dulc[(length(dataupdatedLombardia[,1])-15):length(dataupdatedLombardia[,1]),]),totale_casi=predict(tot.cas.gam, dulc[(length(dataupdatedLombardia[,1])-15):length(dataupdatedLombardia[,1]),]),nuovi_positivi=predict(is.dom.gam,dulc[(length(dataupdatedLombardia[,1])-15):length(dataupdatedLombardia[,1]),]), picco=dulc[100:115,15] ))
dimessi_guariti=predict(dim.guar.gam, dulc)
dimessi_guariti
predictions1<-predict.glm(dulc.lm8, data.frame(terapia_intensiva=dulc[(length(dataupdatedLombardia[,1])-15):length(dataupdatedLombardia[,1]),2],dimessi_guariti=dulc[(length(dataupdatedLombardia[,1])-15):length(dataupdatedLombardia[,1]),8],totale_casi=dulc[(length(dataupdatedLombardia[,1])-15):length(dataupdatedLombardia[,1]),10],nuovi_positivi=dulc[(length(dataupdatedLombardia[,1])-15):length(dataupdatedLombardia[,1]),7],picco=dulc[115:130,15]), interval="confidence", level=.95)                        
predictions1=exp(predictions1)

predictions1 <- data.frame(totale_positivi = as.numeric((predictions1)), days=dulc[(length(dataupdatedLombardia[,1])-15):length(dataupdatedLombardia[,1]),13])
predictions=exp(predictions)
predictions <- data.frame(totale_positivi = as.numeric((predictions)), days=dulc[(length(dataupdatedLombardia[,1])-15):length(dataupdatedLombardia[,1]),13])
predictions
#random_forest <- data.frame(totale_positivi = as.numeric((random_forest[(length(dataupdatedLombardia[,1])-15):length(dataupdatedLombardia[,1])])), days=dulc[(length(dataupdatedLombardia[,1])-15):length(dataupdatedLombardia[,1]),13]) 
#dulc[(length(dataupdatedLombardia[,1])-15):length(dataupdatedLombardia[,1]),5]
  png(file="predictions1.png",width=6, height=4, units="in", res=300)
plot( exp(predictions[1:16,2]),exp(predictions[1:16,1]), col="red",
ylim=c(40000,65000))
points( exp(dulc[(length(dataupdatedLombardia[,1])-15):length(dataupdatedLombardia[,1]),13]), exp(dulc[(length(dataupdatedLombardia[,1])-15):length(dataupdatedLombardia[,1]),5]), col="black",lwd = 1)
points(exp( predictions1[1:16,2]),exp(predictions1[1:16,1]), col="blue")
points( exp(random_forest[1:16,2]),exp(random_forest[1:16,1]), col="green")
dev.off()
data = data.frame(days=dulc[(length(dataupdatedLombardia[,1])-15):length(dataupdatedLombardia[,1]),13], totale_positivi=dulc[(length(dataupdatedLombardia[,1])-15):length(dataupdatedLombardia[,1]),5])

```



1 - 6 novembre bianca
6 - 29 novembre rossa 
29 - 13 dicembre arancione
24 a 27 dicembre rossa 
28-29-30 dicembre arancione
31 - 3 gennaio rossa
4 gennaio arancione
5-6 gennaio rossa
7-8 gennaio gialla
9-15 gennaio arancione
16-23 gennaio rossa
24-1 febbraio arancione
1-ora gialla




By building your model, evaluate the inclusion of some covariates and their effect on the response variable. Some possible covariates could regard: the regional colors (yellow, orange, red), the partial lockdown regime, some region-specific laws and rules, etc.

Check the model fit by using the proper tools, such as residuals plots.

Provide 10-15 days-forward predictions and check their accuracy.

Compare alternative models in terms of predictive information criteria and comment.


